---
title: "BoCo EDA - Andy Pickering"
output: 
  html_document: 
    keep_md: yes
    toc: yes
  html_notebook: 
    toc: yes
---


# Introduction

- This project is looking at patterns of homelessness-related arrests and possible relationship to weather in Boulder County. 
- It was started by Sam Zhang, who gathered and merged the data: <https://github.com/samzhang111/boco-jail-analysis>.
- Data is in a Google drive folder shared by Sam, which I have synced locally on my machine.
- This analysis is in a github repository: <https://github.com/andypicke/boco_homeless>

## Hypotheses To Explore
- How does weather affect daily bookings? weather: *downtown-boulder-weather.csv*. bookings: *bookings-with-transient-status.csv* (you'll need to aggregate this one up to the daily level)
- What proportion of jail bookings for housed/unhoused folks are for homelessness related charges? (by department? by year?) we pulled out charges that came from a list of antihomelessness ordinances compiled by some folks at Denver University, and marked them in the bookings dataset. If any of them were true, we marked the *antihomeless* column as true.




# Weather

- Add more details about where weather data comes from (on Sam's github?)
- What is *SNWD*? Some kind of accumulated snowfall?
- What are *WT01*,*WT02* etc.?
- *SNOW* is missing data from about 2011-204?

```{r Load Libraries}
suppressPackageStartupMessages(library(dplyr))
library(readr)
suppressPackageStartupMessages(library(lubridate))
library(ggplot2)
```

## Load weather data
```{r}
wea <- read_csv('/Users/Andy/Google Drive/boco-jail/downtown-boulder-weather.csv',col_types = cols())
glimpse(wea)
```


```{r}
summary(wea)
```


```{r}
wea$DATE <- lubridate::ymd(wea$DATE)
wea <- wea %>% select( DATE,PRCP,SNOW,TMAX,TMIN)
head(wea)
```


## Plot temperature timeseries
```{r}
wea %>%
        ggplot(aes(DATE,TMAX))+
        geom_point() +
        ylab('Max Temp') +
        ggtitle('Downtown Boulder Weather')
```

## Plot precipiation timeseries
- I assume in inches?
```{r}
wea %>%
        ggplot(aes(DATE,PRCP))+
        geom_point() +
        ylim(0,3)
```


## Plot snow timeseries 
- I assume this is daily snow fall in inches?
- Looks like we are missing a chunk of snowfall data from about 2011-2014 (pretty sure it snowed during those years :) )
```{r}
wea %>%
        ggplot(aes(DATE,SNOW))+
        geom_point() +
        ylim(0,3)
```









# Bookings

Some info on the *bookings-with-transient-status.csv* data:
- (Sam) These are bookings in Boulder County Jail only.
- (Sam) Each row is an individual booking.
- (Sam) `boulder` means the arrest was made by Boulder PD
- (Sam) Column `transient` for whether someone was homeless or not.
- (Sam) I created indicator variables for a variety of antihomeless charges (that a local law school came up with). `antihomeless` is true if any of the antihomeless charges existed in the booking. So one source of error could be that if someone was arrested for an antihomeless charge AND a more serious offense, I still mark it as "antihomeless". 
- (Sam) I don't remember making `any_antihomeless` -- I suspect it's identical to `antihomeless` and introduced by accident
- (Andy) There are two *Booking Time* columns? I will use *booking_time* for now, as it seems better.
- (Andy) What are *location*s? Is that where they were booked?

## Load bookings data
```{r}
bk <- read_csv('/Users/Andy/Google Drive/boco-jail/bookings-with-transient-status.csv',col_types = cols())
glimpse(bk)
```


## Let's clean it up a little
```{r}
bk <- bk %>% select(-`Booking Time`)
names(bk) <- tolower(names(bk))
names(bk) <- gsub(' ','_',names(bk))
bk$arresting_agency <- as.factor(bk$arresting_agency)

to_log <- function(a_col){
        as.logical(a_col)
}
cols_to_log <- c('camping','boulder','urination','vehicle_as_residence','public_obstruct','public_trespass','begging','antihomeless','smoking','any_antihomeless','transient')
bk[cols_to_log] <- lapply(bk[cols_to_log],to_log)

glimpse(bk)
```




### What percent of arrestees are transient?
```{r}

mean(bk$transient,na.rm = TRUE)*100
```

### What percent of arrests are 'antihomeless' (true if any charges were related to 'antihomeless charges')
```{r}

mean(bk$antihomeless)*100
```

### What percent of arrests were made by Boulder PD?
```{r}
mean(bk$boulder)*100
```

###
```{r}
levels(bk$arresting_agency)
```


### What agencies have the most arrests?
```{r}

bk %>%
        group_by(arresting_agency) %>%
        tally() %>%
        arrange( desc(n) ) %>%
        ggplot(aes(x=reorder(arresting_agency,n),y=n))+
        geom_bar(stat='identity',aes(fill=arresting_agency)) +
        coord_flip()
```



### Aggregate Monthly
- I thought there would be more arrests in winter, but there doesn't seem to be that big of a difference.

```{r}
bk$month_ <- lubridate::month(bk$arrest_date,label=TRUE)
bk$year <- lubridate::year(bk$arrest_date)
bk %>% 
        filter(year %in% c(2011,2012,2013,2015,2015)) %>%
        group_by(year,month_) %>%
        tally() %>%
        ggplot(aes(month_,n))+
        geom_point() +
        geom_bar(stat='identity',aes(fill=month_)) +
        facet_wrap(~year)
```


### Aggregate by wkday



### Try aggregating to daily level?
- Looks like daily number of arrests is decreasing over time? 
- Not sure if actual, or something to do with how data is recorded.
```{r}
bk %>% 
        filter(arrest_date>"2000-01-01") %>%
        group_by(arrest_date) %>%
        tally() %>%
        ggplot(aes(arrest_date,n)) +
        geom_point(alpha=0.2) +
        ylim(0,60) +
        geom_smooth(method="lm")
        
```


## Relationship between arrests and weather
- Scatter plot arrests vs weather variables
- Fit linear regressions between arrests and weather
```{r}

```



