---
title: "BoCo EDA - Andy Pickering"
output: 
  html_document: 
    keep_md: yes
    toc: yes
  html_notebook: 
    toc: yes
---


# Introduction

- This project is looking at patterns of homelessness-related arrests and possible relationship to weather in Boulder County. 
- It was started by Sam Zhang, who gathered and merged the data: <https://github.com/samzhang111/boco-jail-analysis>.
- Data is in a Google drive folder shared by Sam, which I have synced locally on my machine.
- This analysis is in a github repository: <https://github.com/andypicke/boco_homeless>

## Hypotheses To Explore
- How does weather affect daily bookings? weather: *downtown-boulder-weather.csv*. bookings: *bookings-with-transient-status.csv* (you'll need to aggregate this one up to the daily level)
- What proportion of jail bookings for housed/unhoused folks are for homelessness related charges? (by department? by year?) we pulled out charges that came from a list of antihomelessness ordinances compiled by some folks at Denver University, and marked them in the bookings dataset. If any of them were true, we marked the *antihomeless* column as true.



```{r Load Libraries}
suppressPackageStartupMessages(library(dplyr))
library(readr)
suppressPackageStartupMessages(library(lubridate))
library(ggplot2)
```





# Bookings

Some info on the *bookings-with-transient-status.csv* data:
- *fta* is "Failure to Appear"
- *ftc* is "Failure to Comply"
- (Sam) These are bookings in Boulder County Jail only.
- (Sam) Each row is an individual booking.
- (Sam) `boulder` means the arrest was made by Boulder PD
- (Sam) Column `transient` for whether someone was homeless or not.
- (Sam) I created indicator variables for a variety of antihomeless charges (that a local law school came up with). `antihomeless` is true if any of the antihomeless charges existed in the booking. So one source of error could be that if someone was arrested for an antihomeless charge AND a more serious offense, I still mark it as "antihomeless". 
- (Sam) I don't remember making `any_antihomeless` -- I suspect it's identical to `antihomeless` and introduced by accident
- (Andy) There are two *Booking Time* columns? I will use *booking_time* for now, as it seems better.
- (Andy) What are *location*s? Is that where they were booked?
- (Andy) *Booked* and *booking_time* appear to be duplicates?

## Load bookings data
```{r}
bk <- read_csv('/Users/Andy/Google Drive/boco-jail/bookings-with-transient-status.csv',col_types = cols())
glimpse(bk)
```


## Let's clean it up a little
```{r}


# 'Booking Time' appears to be a duplicate/bad column
bk <- bk %>% select(-`Booking Time`)

# make columns names lowercase and remove spaces
names(bk) <- tolower(names(bk))
names(bk) <- gsub(' ','_',names(bk))

# some vars make more sense as factors
bk$arresting_agency <- as.factor(bk$arresting_agency)
bk$location <- as.factor(bk$location)
bk$race <- as.factor(bk$race)

# add wkday,month,year so we can aggregate by those
bk$wkday <- lubridate::wday(bk$booking_date,label=TRUE)
bk$month_ <- lubridate::month(bk$arrest_date,label=TRUE)
bk$year <- lubridate::year(bk$arrest_date)

# modern reporting seems to start in 2000 (yearly totals go from less 100 before 2000, to order of 10,000 starting w/ 2000). Keep only 2000 on here.
bk <- bk %>% filter(year>=2000)


bk$booked_date <- lubridate::date(bk$booked)

# a bunch of vars should be logical
to_log <- function(a_col){
        as.logical(a_col)
}
cols_to_log <- c('camping','boulder','urination','vehicle_as_residence','public_obstruct','public_trespass','begging','antihomeless','smoking','any_antihomeless','transient','fta','ftc')
bk[cols_to_log] <- lapply(bk[cols_to_log],to_log)

# clean up some bad age values?
# had some bad values  (ie age=-1 or age=133)
bk$age <- floor(lubridate::interval(bk$dob,bk$arrest_date)/years(1))
bk <- bk %>% filter(age<120 & age>1)


glimpse(bk)
```


### Distribution of ages
```{r}
hist(bk$age)
#summary(bk$age)
```


### What percent of arrestees are transient?
```{r}
mean(bk$transient,na.rm = TRUE)*100
```

### What percent of arrests are 'antihomeless' (true if any charges were related to 'antihomeless charges')
```{r}
mean(bk$antihomeless)*100
```


### What are the arresting agencies?
```{r}
levels(bk$arresting_agency)
```

### What percent of arrests were made by Boulder PD?
```{r}
mean(bk$boulder)*100
```


### What agencies have the most arrests?
```{r}

bk %>%
        group_by(arresting_agency) %>%
        tally() %>%
        arrange( desc(n) ) %>%
        ggplot(aes(x=reorder(arresting_agency,n),y=n))+
        geom_bar(stat='identity',aes(fill=arresting_agency)) +
        coord_flip()
```

### Number of arrests by race
```{r}
bk %>% group_by(race) %>% 
        tally() %>% 
        arrange(desc(n)) %>% 
        ggplot(aes(x=reorder(race,n),y=n))+
        geom_bar(stat='identity',aes(fill=race)) +
        coord_flip()
```


### Aggregate Monthly
- I thought there would be more arrests in winter, but there doesn't seem to be that big of a difference.

```{r}
bk %>% 
        filter(year %in% c(2011,2012,2013,2015,2015)) %>%
        group_by(year,month_) %>%
        tally() %>%
        ggplot(aes(month_,n))+
        geom_point() +
        geom_bar(stat='identity',aes(fill=month_)) +
        facet_wrap(~year)
```

### transients only
```{r}
bk %>% 
        filter(year %in% c(2011,2012,2013,2015,2015)) %>%
        filter(transient==TRUE) %>% 
        group_by(year,month_) %>%
        tally() %>%
        ggplot(aes(month_,n))+
        geom_point() +
        geom_bar(stat='identity',aes(fill=month_)) +
        facet_wrap(~year)
```

### Aggregate by wkday

```{r}
bk %>%
        filter(!is.na(wkday)) %>%
        group_by(wkday) %>%
        tally() %>%
        ggplot( aes(wkday,n)) +
        geom_col(aes(fill=wkday)) +
        ggtitle('Total Arrests By Day, for ALL data')
```


### transients only
```{r}
bk %>%
        filter(!is.na(wkday)) %>%
        filter(transient==TRUE) %>% 
        group_by(wkday) %>%
        tally() %>%
        ggplot( aes(wkday,n)) +
        geom_col(aes(fill=wkday)) +
        ggtitle('Total Arrests By Day, for ALL data, transients only')
```


### By weekday, for separate years
```{r}
bk %>%
        filter(!is.na(wkday)) %>%
        filter(year>1999) %>%
        group_by(year,wkday) %>%
        tally() %>%
        ggplot( aes(wkday,n)) +
        geom_col(aes(fill=wkday)) +
        facet_wrap(~year) +
        ggtitle('Total Arrests By Day, for each year')
```


### By weekday, for separate months
```{r}
bk %>%
        filter(!is.na(wkday)) %>%
        filter(year>1999) %>%
        group_by(month_,wkday) %>%
        tally() %>%
        ggplot( aes(wkday,n)) +
        geom_col(aes(fill=wkday)) +
        facet_wrap(~month_) +
        ggtitle('Total Arrests By Day, for each month, includes all years')
```

### By weekday, for separate months, transients only
```{r}
bk %>%
        filter(!is.na(wkday)) %>%
        filter(transient==TRUE ) %>% 
        filter(year>1999) %>%
        group_by(month_,wkday) %>%
        tally() %>%
        ggplot( aes(wkday,n)) +
        geom_col(aes(fill=wkday)) +
        facet_wrap(~month_) +
        ggtitle('Total Arrests By Day, for each month, includes all years')
```


### Try aggregating to daily level?
- Looks like daily number of non-transient arrests is decreasing over time? 
- daily number of transient arrests is increasing slightly
- Not sure if actual, or something to do with how data is recorded.
```{r}
bk %>% 
        filter(arrest_date>"2000-01-01") %>%
        group_by(arrest_date,transient) %>%
        tally() %>%
        ggplot(aes(arrest_date,n)) +
        geom_point(alpha=0.2,aes(col=transient)) +
        ylim(0,60) +
        geom_smooth(method='lm') +
        facet_wrap(~transient)
```

### Look at number of arrests for 'anti-homeless' charges

```{r}
bk %>% 
        filter(antihomeless==TRUE) %>% 
        group_by(arrest_date) %>% 
        tally() %>% 
        ggplot(aes(arrest_date,n)) +
        geom_hex() +
        geom_smooth(method = 'lm') +
        ggtitle("Daily # arrests for anti homeless charges")
```







# Weather

- Add more details about where weather data comes from (on Sam's github?)
- What is *SNWD*? Some kind of accumulated snowfall?
- What are *WT01*,*WT02* etc.?
- *SNOW* is missing data from about 2011-204?


## Load weather data
```{r}
wea <- read_csv('/Users/Andy/Google Drive/boco-jail/downtown-boulder-weather.csv',col_types = cols())
glimpse(wea)
```


```{r}
summary(wea)
```


```{r}
names(wea) <- tolower(names(wea))
wea$date <- lubridate::ymd(wea$date)
wea <- wea %>% select( date,prcp,snow,tmax,tmin)
head(wea)
```


## Plot temperature timeseries
```{r}
wea %>%
        ggplot(aes(date,tmax))+
        geom_point() +
        ylab('Max Temp') +
        ggtitle('Downtown Boulder Weather')
```

## Plot precipiation timeseries
- I assume in inches?
```{r}
wea %>%
        ggplot(aes(date,prcp))+
        geom_point() +
        ylim(0,3)
```


## Plot snow timeseries 
- I assume this is daily snow fall in inches?
- Looks like we are missing a chunk of snowfall data from about 2011-2014 (pretty sure it snowed during those years :) )
```{r}
wea %>%
        ggplot(aes(date,snow))+
        geom_point() +
        ylim(0,3)
```









# Relationship between arrests and weather

## Let's merge the booking and weather data first

```{r}
bk_wea <- left_join(bk,wea,by=c('arrest_date'='date'))
glimpse(bk_wea)
```

## Scatter plot daily arrests vs weather variables
- There doesn't seem to be much of a relationship between daily # arrests and temperature. 
```{r}
bk %>%
        group_by(arrest_date) %>% 
        tally() %>% 
        left_join(wea,by=c('arrest_date'='date')) %>%
        ggplot(aes(tmin,n)) +
        geom_hex() +
        geom_smooth(method = 'lm') +
        ggtitle("Daily # arrests vs min Temp")
```



## Scatter plot daily _transient_ arrests vs weather variables
- Not too much for transient only either.
```{r}
bk%>%
        filter(transient==TRUE) %>% 
        group_by(arrest_date) %>% 
        tally() %>% 
        left_join(wea,by=c('arrest_date'='date')) %>%
        ggplot(aes(tmin,n)) +
        geom_hex() +
        geom_smooth(method = 'lm') +
        ggtitle("Daily # *transient* arrests vs min Temp.")
```


## Scatter plot daily _antihomeless_ arrests vs weather variables
- Shows a slight positive trend, which is opposite what I would expect (thought there would be more arrests in colder weather)
```{r}
bk%>%
        filter(antihomeless==TRUE) %>% 
        group_by(arrest_date) %>% 
        tally() %>% 
        left_join(wea,by=c('arrest_date'='date')) %>%
        ggplot(aes(tmin,n)) +
        geom_hex() +
        geom_smooth(method = 'lm') +
        ggtitle("Daily anti-homeless arrests vs min temp")
```

## Try a logistic regression of antihomeless arrest as function of temperature, precip, etc.?


## Look at individual years?
```{r}
bk%>%
        filter(antihomeless==TRUE) %>% 
        group_by(arrest_date) %>% 
        tally() %>% 
        left_join(wea,by=c('arrest_date'='date')) %>%
        mutate(year=lubridate::year(arrest_date)) %>% 
        ggplot(aes(tmin,n)) +
        geom_hex() +
        geom_smooth(method = 'lm') +
        facet_wrap(~year)
#        ggtitle("Daily anti-homeless arrests vs min temp")

```



